---
output: html_document
---
COURSERA-REPORODUCIBLE RESEARCH PA-1
====================================

**Loading and preprocessing the data**

**1.Loading the data**

The file containing the activity monitoring data is downloaded and stored in current working directory. It is read in R using read.csv function.

```{r,echo=TRUE}
activity_data<-read.csv(".\\activity.csv")
```

**2.Pre-procesing the data**
```{r,echo=TRUE}
str(activity_data)
```

It can be seen that class of *date* column is factor.It is changed to date class using **as.Date** function.

```{r,echo=TRUE}
activity_data$date<-as.Date(activity_data$date)
```


**What is mean total number of steps taken per day?**

**1.Total number of steps taken per day**

Total number of steps taken per day is calculated as
```{r,echo=TRUE}
library(dplyr,warn.conflicts = FALSE)
steps_by_day<-activity_data %>% group_by(date) %>% summarise(Total_steps=sum(steps,na.rm = TRUE))
head(steps_by_day)
```

**2.Histogram of the total number of steps taken each day**

The histogram of total number of steps taken per day is drawn using **plot** function.

```{r,echo=TRUE}
plot(activity_data$date,activity_data$steps,xlab="Date",ylab="No. of Steps per day",main="Histogram of no. of steps taken per day",type="h")
```


**3.Mean and Median of the total number of steps taken per day**

Mean and Median of number of steps taken per day is found using *mean* and *median* function.
```{r,echo=TRUE}
mean_steps<-activity_data %>% group_by(date) %>% summarise(mean_no_steps=mean(steps,na.rm = TRUE))
head(mean_steps)
median_steps<-activity_data %>% group_by(date) %>% summarise(median_no_steps=median(steps))                                                        
head(median_steps)
```

**Average daily activity pattern**

**1.Time-seies plot **

Average number of steps taken per 5-minute interval can be found using dplyr and stoerd in *interval_avg* dataframe.
```{r,echo=TRUE}

interval_avg<-activity_data %>% group_by(interval) %>% summarise(mean_steps=mean(steps,na.rm=TRUE))
```
The time-series plot of average number of steps taken per 5-minute interval is shown below
```{r,echo=TRUE}
plot(interval_avg$interval,interval_avg$mean_steps,type="l",xlab="5-min interval",ylab="avg no. of steps",main="Avg no. of steps for every 5-min interval")
```

**2.calculating 5-minute interval with maximum number of steps**

Data Frame *interval_avg* contains the average number of steps taken for every 5-minute period.From that it can be found that time period containing the maximum value for average number of steps.
```{r,echo=TRUE}
max_steps_time_period<-interval_avg[which.max(interval_avg$mean_steps),1]
```

It can be found that time period `r max_steps_time_period` contains the maximum value for average number of steps.

**Differences in activity patterns between weekdays and weekend **

**1.creating new factor variable to differntiate weekday and weekend **

Using *weekdays* function,day of the week can be found from the date.A new column containing day of the week is added to original dataframe.
```{r,echo=TRUE}
activity_data$weekday<-weekdays(activity_data$date)
```

**2.Creating plot seperately for weekdays and weekends **

This weekday column is used for filtering(with *dplyr*) the weekdays and weekend days seperately and average number of steps taken are calculated seperately.
```{r,echo=TRUE}
weekend_avg<-filter(activity_data,grepl('Saturday|Sunday',weekday)) %>% group_by(interval) %>% summarise(mean_steps=mean(steps,na.rm=TRUE))
weekdays_avg<-filter(activity_data,!grepl('Saturday|Sunday',weekday)) %>% group_by(interval) %>% summarise(mean_steps=mean(steps,na.rm=TRUE))
```
Plots for average number of steps taken for each 5-min interval is made seperately for weekdays and weekend days.
```{r,echo=TRUE}
par(mfrow=c(2,1))
plot(weekdays_avg$interval,weekdays_avg$mean_steps,type="l",xlab="5-min interval",ylab="avg no.of steps",sub="weekdays plot")
plot(weekend_avg$interval,weekend_avg$mean_steps,type="l",xlab="5-min interval",ylab="avg no. of steps",sub="weekend plot")
```

**Dealing with missing values**

**1.calculating the total number of missing values **

Total number of missing values in the given data is calculated as below.

```{r,echo=TRUE}
NA_count<-length(activity_data[is.na(activity_data$steps),1])
```
It is found that `r NA_count` rows are having NA in steps column.
