COURSERA-REPORODUCIBLE RESEARCH PA-1
====================================

The file containing the activity monitoring data is read in R using read.csv function.

```{r,echo=TRUE}
activity_data<-read.csv(".\\activity.csv")
```
It can be seen that class of *date* column is factor.It is changed to date class using **as.Date** function.

```{r,echo=TRUE}
str(activity_data)
activity_data$date<-as.Date(activity_data$date)
```
The histogram of total number of steps taken per day is drawn using **plot** function.

```{r,echo=TRUE}
plot(activity_data$date,activity_data$steps,xlab="Date",ylab="No. of Steps per day",main="Histogram of no. of steps taken per day",type="h")
```



Mean and Median of number of steps in the whole dataset is foud using *mean* and *median* function.

```{r,echo=TRUE}
steps_mean<-mean(activity_data$steps,na.rm=TRUE)
steps_median<-median(activity_data$steps,na.rm=TRUE)
steps_mean
steps_median
```

Simlarly average number of steps taken per 5-minute interval can be found using dplyr.
```{r,echo=TRUE}
library(dplyr)
interval_avg<-activity_data %>% group_by(interval) %>% summarise(mean_steps=mean(steps,na.rm=TRUE))
```
The time-series plot of average number of steps taken per 5-minute interval is shown below
```{r,echo=TRUE}
plot(interval_avg$interval,interval_avg$mean_steps,type="l",xlab="5-min interval",ylab="avg no. of steps",main="Avg no. of steps for every 5-min interval")
```

Data Frame *interval_avg* contains the average number of steps taken for every 5-minute period.From that it can be found that **time period 835** contains the maximum value for average number of steps.
```{r,echo=TRUE}
max_steps_time_period<-interval_avg[which.max(interval_avg$mean_steps),1]
max_steps_time_period
```
Using *weekdays* function,day os the week can be found from the date.A new column containing day of the week is added to original dataframe.
```{r,echo=TRUE}
activity_data$weekday<-weekdays(activity_data$date)
```
This weekday column is used for filtering(with *dplyr*) the weekdays and weekend days seperately and average number of steps taken are calculated seperately.
```{r,echo=TRUE}
weekend_avg<-filter(activity_data,grepl('Saturday|Sunday',weekday)) %>% group_by(interval) %>% summarise(mean_steps=mean(steps,na.rm=TRUE))
weekdays_avg<-filter(activity_data,!grepl('Saturday|Sunday',weekday)) %>% group_by(interval) %>% summarise(mean_steps=mean(steps,na.rm=TRUE))
```
Plots for average number of steps taken for each 5-min interval is made seperately for weekdays and weekend days.
```{r,echo=TRUE}
par(mfrow=c(2,1))
plot(weekdays_avg$interval,weekdays_avg$mean_steps,type="l",xlab="5-min interval",ylab="avg no.of steps",sub="weekdays plot")
plot(weekend_avg$interval,weekend_avg$mean_steps,type="l",xlab="5-min interval",ylab="avg no. of steps",sub="weekend plot")
```